\textbf{Are regime changes always bad economics? Evidence from daily financial data}

### 2019-04-28 doc/paper.tex
Update intro

The image of a coup or popular uprising ousting an autocratic ruler and installing a democratic government is an evocative one. However, both recent events and prior research suggest that coups rarely lead to democracy. In the wake of the Arab Spring - in which rulers were deposed in four countries - only Tunisia ultimately transitioned to democracy. These results are in line with prior research that claims that in autocracies coups rarely lead to democratization, and in democracies they cause instability \citep{derpanopoulos2015coups, marinov2014coups, powell2011global, thyne2016coup, varol2011democratic}.

Of course, coups and other regime changes not only impact a country's political environment but its economy as well. For instance, investors, multinational firms, development agencies, and aid organizations, rank political risk as a top consideration when making investment decisions in emerging markets.\footnote{In the 2013 Multilateral Investment Guarantee Agency (MIGA) \citet{wipr2013} report (the last report published), executives of multinational enterprises (MNEs) ranked political risk as the second most important constraint for foreign direct investment (FDI) in developing countries over the next three years (after macroeconomic instability). In \citet{wipr2011} and  \citet{wipr2012}, political risk was ranked the most important constraint for FDI in developing countries - even greater than macroeconomic instability. The types of political risk of most concern to investors in developing countries (ranked in order of importance) were 1) adverse regulatory changes, 2) breaches of contract, 3) transfers and convertibility restrictions, 4) civil disturbances, 5) non-honoring of government guarantees, 6) expropriation nationalization, 7) terrorism and 8) war.} Common rationales are that unstable governments act myopically and adopt inefficient policies \citep{svensson1998investment,devereux1998political}\footnote{\citet{svensson1998investment} argues that governments are less likely to invest in the legal system and the protection of property rights when political uncertainty is high. Similarly, \citet{devereux1998political} propose a model in which political instability causes governments to leave fewer assets to their successors which forces them into increasing capital taxes. The knowledge of future taxation then causes the private sector to reduce current investment which reduces future output.} and that policy uncertainty depresses business activity \citep{baker2016measuring}. Such sentiments are supported by empirical research that suggests political instability---as measured by regime change frequency or political violence---is negatively correlated with investment, financial development, and GDP growth in cross country regressions \citep{aisen2013does, alesina1996income, alesina1996political, fosu1992political, jong2009measurement, roe2011political} and that the variance of stock market returns increases in response to economic policy uncertainty \citep{leblang2005government, jensen2005market, liu2015economic}.

However, such hypotheses are often overly simplistic. Not all political instability is equivalent - resignations, assassinations, and coups are all examples of regime changes, but may have disparate effects on financial returns. Resignations typically occur when an unpopular leader steps down and may signal the coming of a more effective leader. Assassinations are not always related to the effectiveness of a leader, can occur seemingly at random, and the successor to the assassinated leader may be unclear. Coups can occur in the name of democracy or autocracy, and in general, signal relative weakness in the current political system. Further, political instability may be seen as a positive event if the current regime is regarded as strongly anti-business or anti-global.  \citet{alesina1996political} take the possibility of a new government adopting better economic policies seriously, but argue that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions. Likewise, \citet{londregan1990poverty}, find that coups have no impact on economic growth and hypothesize that this may because there is a bimodal distribution of coups, some of which enhance growth and others which restrict it. \citet{meyersson2016political} provides  support for this hypothesis, finding that coup attempts that occur in democratic regimes are associated with negative GDP growth, while those that occur in autocratic regimes are associated with positive GDP growth.

In this study, we examine the impact of ``irregular'' regime changes and public protests on economic performance and remain agnostic about the direction of the effect. In contrast to the aforementioned cross country studies, we  conduct our analysis using daily financial data. Market expectations in the form of stock market returns are therefore used as an indicator of whether the market views different types of potentially destabilizing political events as ``good'' for economic growth.  The advantage of this event study approach is that it allows us to use time-series variation in stock returns to estimate treatment effects and mitigate the endogeneity problems that are nearly ubiquitous in cross country regressions.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.][]{fisman2001estimating,faccio2006politically,goldman2009politically}. Studies in ``forensic economics" have used abnormal returns from political events to locate or examine transactions such as insider trading \citep{dube2011coups}, illegal arms trading \citep{dellavigna2010detecting}, and the impact of hostilities on the financial performance of diamond mining firms in Angola \citep{guidolin2007diamonds}.}

We find that financial volatility increases dramatically following (and just before) coups d'etat, assassinations, and resignations. However, unlike the previously mentioned studies, we find that these irregular regime changes have disparate effects on the direction of stock returns. Abnormal returns following resignations are large and positive (4\%) while abnormal returns following assassinations are negative and smaller in magnitude (2\%). The impact of coups tends to be negative (2\%), but some events result in positive abnormal returns of 10\% or more.

These findings are consistent with research that shows that ``good coups'' - coups that lead to democratization or economic liberalization - are not the norm and therefore on average coups should not be expected to lead to increased economic development and positive market returns \citep{derpanopoulos2015coups, marinov2014coups, powell2011global, thyne2016coup, varol2011democratic}. However, while ``good coups'' may be the minority, regime changes that attempt to overthrow anti-business autocrats may be expected to result in positive abnormal returns as investors see little risk of a replacement government ``worse'' than the status quo.

Our main finding is therefore that the type of political event and its expected impact on economic policy determines the direction of abnormal returns. Events expected to lead to more stable governance, economic liberalization, or democratization (such as willful resignations and coups that overthrow protectionist or leftist autocrats) are associated with positive returns, while those that consolidate authoritarian rule, exacerbate poor economic policies, or merely increase policy uncertainty have the opposite effect.

@@

\section{Conclusion}
This paper has examined the effects of irregular regime changes on financial returns. It uses an event study approach to show that investors expect irregular regime changes to have large effects on equity returns. This methodology is less susceptible to endogeneity biases than studies that use cross country data.

>>>>

The image of a coup or popular uprising ousting an autocratic ruler and installing a democratic government is an evocative one. However, both recent events and prior research suggest that coups rarely lead to democracy. In the wake of the Arab Spring---in which rulers were deposed in four countries---only Tunisia ultimately transitioned to democracy. These results are in line with prior research that claims that in autocracies coups rarely lead to democratization, and in democracies they cause instability \citep{derpanopoulos2015coups, marinov2014coups, powell2011global, thyne2016coup, varol2011democratic}.

Of course, coups and other regime changes not only impact a country's political environment but its economy as well. For instance, investors, multinational firms, development agencies, and aid organizations rank political risk as a top consideration when making investment decisions in emerging markets.\footnote{In the 2013 Multilateral Investment Guarantee Agency (MIGA) \citet{wipr2013} report (the last report published), executives of multinational enterprises (MNEs) ranked political risk as the second most important constraint for foreign direct investment (FDI) in developing countries over the next three years (after macroeconomic instability). In \citet{wipr2011} and  \citet{wipr2012}, political risk was ranked the most important constraint for FDI in developing countries - even greater than macroeconomic instability. The types of political risk of most concern to investors in developing countries (ranked in order of importance) were 1) adverse regulatory changes, 2) breaches of contract, 3) transfers and convertibility restrictions, 4) civil disturbances, 5) non-honoring of government guarantees, 6) expropriation nationalization, 7) terrorism and 8) war.} Common rationales are that unstable governments act myopically and adopt inefficient policies \citep{svensson1998investment,devereux1998political}\footnote{\citet{svensson1998investment} argues that governments are less likely to invest in the legal system and the protection of property rights when political uncertainty is high. Similarly, \citet{devereux1998political} propose a model in which political instability causes governments to leave fewer assets to their successors which forces them into increasing capital taxes. The knowledge of future taxation then causes the private sector to reduce current investment which reduces future output.} and that policy uncertainty depresses business activity \citep{baker2016measuring}. Such sentiments are supported by empirical research that suggests political instability---as measured by regime change frequency or political violence---is negatively correlated with investment, financial development, and GDP growth in cross country regressions \citep{aisen2013does, alesina1996income, alesina1996political, fosu1992political, jong2009measurement, roe2011political} and that the variance of stock market returns increases in response to economic policy uncertainty \citep{leblang2005government, jensen2005market, liu2015economic}.

But such hypotheses are often overly simplistic. Not all political instability is equivalent - resignations, assassinations, and coups are all examples of regime changes, but may have disparate effects on financial returns. Resignations typically occur when an unpopular leader steps down and may signal the coming of a more effective leader. Assassinations are not always related to the effectiveness of a leader, can occur seemingly at random, and the successor to the assassinated leader may be unclear. Coups can occur in the name of democracy or autocracy, and in general, signal relative weakness in the current political system. Further, political instability may be seen as a positive event if the current regime is regarded as strongly anti-business or anti-global.  \citet{alesina1996political} take the possibility of a new government adopting better economic policies seriously, but argue that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions. Likewise, \citet{londregan1990poverty} find that coups have no impact on economic growth and hypothesize that this may because there is a bimodal distribution of coups, some of which enhance growth and others which restrict it. More recent studies provide support for this hypothesis. \citet{meyersson2016political} finds that coup attempts that occur in democratic regimes are associated with negative GDP growth, while those that occur in autocratic regimes are associated with positive GDP growth. \citet{girardi2018institution} show that Salvador Allende's socialist government was associated with large declines in stock prices while the coup that replaced him boosted them.

In this study, we examine the impact of ``irregular'' regime changes and public protests on economic performance and remain agnostic about the direction of the effect. In contrast to the aforementioned cross country studies but like \citet{girardi2018institution} we  conduct our analysis using daily financial data. Market expectations in the form of stock market returns are used as an indicator of whether the market views different types of potentially destabilizing political events as ``good'' for economic growth. The advantage of this event study approach is that it allows us to use time-series variation in stock returns to estimate treatment effects and mitigate the endogeneity problems that are nearly ubiquitous in cross country regressions.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.][]{fisman2001estimating,faccio2006politically,goldman2009politically}. Studies in ``forensic economics" have used abnormal returns from political events to locate or examine transactions such as insider trading \citep{dube2011coups}, illegal arms trading \citep{dellavigna2010detecting}, and the impact of hostilities on the financial performance of diamond mining firms in Angola \citep{guidolin2007diamonds}.} We analyze 13 coups, 8 assassinations, 15 forced resignations, 11 public protests, and 1 natural experiment provided by the failed 2002 coup in Venezuela. Our study is therefore not only able to use an event-study approach to enhance the reliability of the estimates, but is not limited to a single case study as in \citet{girardi2018institution}. The relatively large sample enhances external validity and allows us to show that different types of political events have disparate economic effects.

We find that coups, assassinations, resignations, and public protests cause large increases in financial volatility. Abnormal returns following resignations are large and positive (4\%) while abnormal returns following assassinations are negative and smaller in magnitude (2\%). The average impact of coups is negative (2\%); however, stock prices increased dramatically during the failed overthrow of Hugo Chavez's socialist regime and decreased by approximately the same magnitude following his return to leadership. The effect of public protests on mean stock returns is uncertain, with some protests such as those during the Egpytian Revolution depressing them and others have the opposite effect.

These findings are consistent with research that shows that ``good coups''---coups that lead to democratization or economic liberalization---are not the norm and therefore on average coups should not be expected to lead to increased economic development and positive market returns \citep{derpanopoulos2015coups, marinov2014coups, powell2011global, thyne2016coup, varol2011democratic}. However, while ``good coups'' may be the minority, regime changes that attempt to overthrow anti-business autocrats may be expected to result in positive abnormal returns as investors see little risk of a replacement government ``worse'' than the status quo.

Our main conclusion is therefore that the type of political event and its expected impact on economic policy determines the direction of abnormal returns. Events expected to lead to more stable governance, economic liberalization, or democratization (such as willful resignations and coups that overthrow protectionist or leftist autocrats) are associated with positive returns, while those that consolidate authoritarian rule, exacerbate poor economic policies, or merely increase policy uncertainty have the opposite effect.

@@

\section{Conclusion and Discussion}
[Abundance of bad leaders. Coups are still commonplace]

[How we tackle the problem. Strength of our design]

[What our analysis suggest. Summary of findings]

[Compare results to other literature. i.e., magnitude of stock changes. Democratization is not guarantee of positive stock returns]

[Discuss potential mechanisms. Note as a limitation and an avenue for future research.]

This paper has examined the effects of irregular regime changes on financial returns. It uses an event study approach to show that investors expect irregular regime changes to have large effects on equity returns. This methodology is less susceptible to endogeneity biases than studies that use cross country data.
